Study 2 Analysis

### 2024-11-04 misc/discrep.qmd
compute study 2 discrepancy counts

s1_discrep_ids

@@

@@

  # Massachusetts values (Davis Family)
  "Massachusetts", "Cooling", 419, 0, 322,
  "Massachusetts", "Heating", 26751, 27000, 19108,
  "Massachusetts", "Water Heating", 10543, 11000, 5070,
  "Massachusetts", "Refrigerator", 1230, 1000, 1025,
  "Massachusetts", "Other Appliance", 7350, 7000, 6682,
  "Massachusetts", "Total kWh", 46293, 46000, 32207
) %>%
  rowwise() %>%
  mutate(
    state_abrv = case_when(
      state == "Texas" ~ "TEX",
      state == "California" ~ "CAL",
      state == "Colorado" ~ "COL",
      state == "Massachusetts" ~ "MASS"
    ),
    family_name = case_when(
      state == "Texas" ~ "Smith Family",
      state == "California" ~ "Adams Family",
      state == "Colorado" ~ "Wells Family",
      state == "Massachusetts" ~ "Davis Family"
    ))

@@

>>>>

s1_discrep_ids
#saveRDS(s1_discrep_ids,here::here("data/s1_discrep_ids.rds"))

@@

## Study 2

### Counts per unique value of family prior year, and state average

@@

  # Massachusetts values (Davis Family)
  "Massachusetts", "Cooling", 419, 0, 322,
  "Massachusetts", "Heating", 26751, 27000, 19108,
  "Massachusetts", "Water Heating", 10543, 11000, 5070,
  "Massachusetts", "Refrigerator", 1230, 1000, 1025,
  "Massachusetts", "Other Appliance", 7350, 7000, 6682,
  "Massachusetts", "Total kWh", 46293, 46000, 32207
) %>%
  rowwise() %>%
  mutate(
    state_abrv = case_when(
      state == "Texas" ~ "TEX",
      state == "California" ~ "CAL",
      state == "Colorado" ~ "COL",
      state == "Massachusetts" ~ "MASS"
    ),
    app_abrv = case_when(
      appliance == "Cooling" ~ "AC",
      appliance == "Heating" ~ "HEAT",
      appliance == "Water Heating" ~ "WATER",
      appliance == "Refrigerator" ~ "REFRIG",
      appliance == "Other Appliance" ~ "OTHER",
      appliance == "Total kWh" ~ "TOTAL"
    ),
    family_name = case_when(
      state == "Texas" ~ "Smith Family",
      state == "California" ~ "Adams Family",
      state == "Colorado" ~ "Wells Family",
      state == "Massachusetts" ~ "Davis Family"
    ))

@@

### Discrepant Subjects

s2_long %>%
  filter(plan == "plan1",
  id !="s2_208") %>% # s2_208 is just missing a value for a total column - but has all their app values
  group_by(appliance, rounded, state, pct_goal) %>%
  mutate(
    distinct_families = n_distinct(family),
    distinct_state_avg = n_distinct(state_avg),
  ) %>% 
  filter(distinct_families > 1 | distinct_state_avg > 1) %>%
  group_by(appliance, rounded, state, pct_goal, family) %>%
  mutate(family_count = n()) %>%
  group_by(appliance, rounded, state, pct_goal, state_avg) %>%
  mutate(state_avg_count = n()) %>%
  distinct(rounded, state, refClass, appliance, family, state_avg, 
          state_avg_count, family_count) %>%
  mutate(
    # Check family values
    fam_in = map2_lgl(rounded, family, ~ {
      if (.x == "Rounded") {
        .y %in% correct_values_s2$family_rounded
      } else {
        .y %in% correct_values_s2$family_exact
      }
    }),
    # Check state averages
    state_in = pmap_lgl(list(state, appliance, state_avg), function(s, a, sa) {
      correct_row <- correct_values_s2 %>%
        filter(state == s, app_abrv == a)
        #print(nrow(correct_row))  
      if(nrow(correct_row) == 0) FALSE else sa == correct_row$state_avg
    })
  ) %>%
  arrange(rounded, refClass, state, appliance, family_count) %>%
  relocate(rounded, pct_goal, refClass, state, appliance, family, 
          family_count, state_avg, state_avg_count, fam_in, state_in) |> kable()

discrepant_subjects_s2 <- s2_long |> 
  filter(plan == "plan1") |>
  group_by(id, rounded, refClass) |> 
  mutate(
    # Compute state_in based on exact match for state, appliance abbreviation, and state_avg
    state_in = pmap_lgl(list(state, appliance, state_avg), function(s, a, sa) {
      correct_row <- correct_values_s2 %>%
        filter(state == s, app_abrv == a)
      if (nrow(correct_row) == 0) FALSE else sa == correct_row$state_avg
    }),
    # Compute fam_in based on `rounded` value and appropriate family column in `correct_values_s2`
    fam_in = pmap_lgl(list(state, appliance, family, rounded), function(s, a, fam, rnd) {
      correct_row <- correct_values_s2 %>%
        filter(state == s, app_abrv == a)
      if (nrow(correct_row) == 0) {
        FALSE
      } else if (rnd == "Rounded") {
        fam == correct_row$family_rounded
      } else {
        fam == correct_row$family_exact
      }
    })
  )

discrepant_subjects_s2 |> filter(state_in==FALSE | fam_in==FALSE) |>  arrange(state, appliance,refClass,pct_goal,rounded) |> select(-calc,-edu, -plan,-state_dif) |> kable()

s2_discrep_ids <- discrepant_subjects_s2 |> filter(state_in==FALSE | fam_in==FALSE) |> pull(id) |> unique()
s2_discrep_ids
# saveRDS(s2_discrep_ids,here::here("data/s2_discrep_ids.rds"))

# mutate(
#   fam_in = if_else(rounded == "Rounded",
#                   family %in% correct_values_s2$family_rounded,
#                   family %in% correct_values_s2$family_exact),
# # Use `map_lgl` to evaluate each row individually
#   state_in = map_lgl(
#     seq_along(appliance),
#     ~ any(correct_values_s2$appliance[correct_values_s2$state_avg == state_avg[.x]] == appliance[.x])
#   ) )

# 3 s2_185     2 USD      HEAT      10%      Rounded Colorado Calculator    Some graduate school   0.1  18000     16411 plan1 17000  -1000       589
# 4 s2_196     2 USD      HEAT      10%      Rounded Colorado Calculator    College degree         0.1  18000        NA plan1 17000  -1000        NA
# 5 s2_197     2 USD      HEAT      10%      Rounded Colorado Calculator    Some college           0.1  18000     16411 plan1 17000  -1000       589

#  s2 |>
#   mutate(id=paste0("s2_",row_number())) |> relocate(id) |> filter(id=="s2_196") |> 
#    select(-starts_with("AIN")) |> select(starts_with("HEAT"))

# s2 |> filter(id=="s2_196") |> select(starts_with("HEAT"))

s2_long |> filter(state=="Colorado",state_avg==1142,appliance=="WATER")

haven::read_sav(here::here("data/Frequency & Probability Study 2 - 3-24-19.sav")) |>
  mutate(id=paste0("s2_",row_number())) |> 
  relocate(id) |> filter(id=="s2_196") |> 
   select(-starts_with("AIN")) |> select(starts_with("HEAT"))

#s2 |> filter(id=="s2_208") |> select(starts_with("T"))
s2_long |> filter_all(any_vars(is.na(.)))

### 2024-11-04 study2.qmd
compute study 2 discrepancy counts

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
  mutate(id=paste0("s2_",row_number())) |> relocate(id)
s2 <- s2 |> filter(!(id=="s2_78")) # NA for some responses
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

>>>>

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
  mutate(id=paste0("s2_",row_number())) |> relocate(id)
s2 <- s2 |> filter(!(id=="s2_78")) |> # NA for some responses
  filter(!(id=="s2_196")) # NA for state_avg for COL HEAT - (HEATSTATE2)
  # filter(!(id=="s2_208")) # NA for response for California not rounded Total (TOTALAPONE1 and TOTALAPTWO1) - leave them in since all the acutal app values are okay
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

# s2 |> filter(id=="s2_208") |> select(starts_with("T"))
# s2_long |> filter_all(any_vars(is.na(.)))
